The low-frequency Murchison radio environment:\\interference analysis and mitigation

### 2014-10-03 articles/2014/MWA-Radio-Environment/mwa-radio-environment.tex
Work on article

\subsection{The AOFlagger}

\begin{figure}
\begin{center}\hspace*{-0.8cm}\includegraphics[width=10.5cm]{img/bandpass}
\caption{Correlator output RMS over frequency in an MWA high-frequency observation, calculated over all cross-correlated baselines and 112 seconds of data. In this band, the band-pass shows two large discontinuities over frequency, because the MWA receivers apply different digital gains at different frequencies to minimize quantization noise. TODO explain subband bandpass somewhere %TODO
}
\label{fig:bandpass}
\end{center}
\end{figure}

@@

\begin{table*}
\caption{List of observations used in the analyses. The RFI column contains the fraction of visibilities that the initial \textsc{aoflagger} run has classified as RFI (which includes false positives). The DVB column displays the fraction of visibilities that were unusable because of the presence of DVB signals, and which the flagger does not flag. Because the DVB frequencies are only observed in GLEAM observations, this fraction is only given for those. }\label{tbl:obs-list}
\begin{tabular}{|l|l|r|r|r|c|}
\hline
\textbf{Project} &\textbf{Date} & \textbf{Frequencies (MHz)} & \textbf{Duration} & RFI & DVB \\
\hline
EoR high& 2013-08-23 & 167.0--197.7              & 3h & 0.53\% & ---\\
GLEAM & 2013-08-25 & 72.3--133.8, 138.9--230.8 & 7h & 0.94\% & 0\% \\
EoR low& 2013-08-26 & 138.9--169.6              & 6h & 0.81\% & ---\\
GLEAM & 2013-11-05 & 72.3--133.8, 138.9--230.8 & 7h & 1.27\% & 0\% \\
GLEAM & 2013-11-25 & 72.3--133.8, 138.9--230.8 & 7h & 0.69\% & 0\% \\
EoR high& 2014-02-05 & 167.0--197.7              & 6h & 0.54\% & -- \\
GLEAM & 2014-03-16 & 72.3--133.8, 138.9--230.8 & 7h & 0.79\% & 0\% \\
GLEAM & 2014-03-17 & 72.3--133.8, 138.9--230.8 & 7h & 1.64\% & 3\% \\
EoR high& 2014-04-10 & 167.0--197.7              & 6h & 0.68\% & ---\\
GLEAM & 2014-06-18 & 72.3--133.8, 138.9--230.8 & 7h & 0.98\% & 0\% \\
RFI test&2014-08-27& 131.2--161.9     & 2$\times$8 m& 1.85\% & --- \\
\hline
\multicolumn{3}{|r|}{\textbf{Total}} & \textbf{63h} & \textbf{0.96\%} & \textbf{0.5\%}\\
\hline
\end{tabular}
\end{table*}

@@

\section{Detection results}
The second last column of Table~\ref{tbl:obs-list} list for each of the observation nights the ratio of samples that were classified as RFI by the \textsc{aoflagger}.

@@

\noindent\begin{figure*}
\subcaptionbox{RFI contamination found in the 2 m amateur band (146 MHz).}[\linewidth]{%
% obs data: 1061583576, 2013-08-26.
\includegraphics[width=9.1cm]{img/amateur-2m-band-example}\includegraphics[width=9.1cm]{img/amateur-2m-band-flagged}
\label{fig:amateur-2m}
}\\\vspace{3mm}%
\subcaptionbox{Short RFI burst of 2 s centred on 150.17 MHz. This shows that flagging at high time and frequency resolution is necessary to accurately detect RFI.}[\linewidth]{%
% obs data: 1061583576, 2013-08-26.
\includegraphics[width=9.1cm]{img/150_2_mhz_example}\includegraphics[width=9.1cm]{img/150_2_mhz_flagged}
\label{fig:150_2}
}\\\vspace{3mm}%
\subcaptionbox{A weakly-observed transmitter centred on 156.66 MHz. This same frequency is flagged in several correlations, and it can therefore be assumed this is real RFI. This is an example in which the AOFlagger detects an event that would not be detected by regular thresholding. }[\linewidth]{%
% obs data: 1061583576, 2013-08-26.
\includegraphics[width=9.1cm]{img/156_7_mhz_example}\includegraphics[width=9.1cm]{img/156_7_mhz_flagged}
\label{fig:156_7}
}%
\caption{Examples of RFI events found in the EoR low band (138.9--197.7 MHz) in a 2-min snapshot with relatively high RFI contamination. These panels show the Stokes I amplitudes. In the right figures, the result of RFI detection is shown with purple. The horizontally flagged lines are flagged because they are 1.28-MHz subband edge or centre channels, which are unusable because of aliasing of the poly-phase filter bank and DC offsets, respectively.}
\end{figure*}

\noindent\begin{figure}
% obs data: 1061583576, 2013-08-26.
\begin{center}\hspace*{-0.2cm}\includegraphics[width=9.1cm]{img/dvb_example}
\caption{RFI from Digital Video Broadcasting (DVB). Because of its broadband nature, this kind of RFI is not well flagged in the initial AOFlagger step, but is detectable in the global statistics. This type of RFI makes most of the 30.72~MHz bandwidth unusable. It is visible for about 2 h in 1 out of the 6 nights that observed this band. }
\label{fig:dvb_example}
\end{center}
\end{figure}

@@

\noindent\begin{figure*}
\begin{center}\hspace*{-0.2cm}\includegraphics[width=18cm]{img/rfi-per-band}
\caption{RFI per subband}
\label{fig:rfi-per-band}
\end{center}
\end{figure*}

>>>>

\subsection{The \textsc{aoflagger} RFI detector}

\begin{figure}
\begin{center}\hspace*{-0.8cm}\includegraphics[width=10.5cm]{img/bandpass}
\caption{Correlator output RMS over frequency in an MWA high-frequency observation, calculated over all cross-correlated baselines and 112 seconds of data. In this band, the band-pass shows two large discontinuities over frequency, because the MWA receivers apply different digital gains at different frequencies to minimize quantization noise. The 1.28~MHz pattern is due to the imperfect poly-phase filter.
}
\label{fig:bandpass}
\end{center}
\end{figure}

@@

\begin{table*}
\caption{List of observations used in the analyses. The RFI column contains the fraction of visibilities that the initial \textsc{aoflagger} run has classified as RFI (which includes false positives). The DVB column displays the fraction of visibilities that were unusable because of the presence of DVB signals, and which the flagger does not flag. Not all observations cover the DVB frequencies. }\label{tbl:obs-list}
\begin{center}%
\begin{tabular}{|l|l|r|r|r|c|}
\hline
\textbf{Project} &\textbf{Date} & \textbf{Frequencies (MHz)} & \textbf{Duration} & RFI & DVB \\
\hline
EoR high& 2013-08-23 & 167.0--197.7              & 3h & 0.53\% & 0\% \\
GLEAM & 2013-08-25 & 72.3--133.8, 138.9--230.8 & 7h & 0.94\% & 0\% \\
EoR low& 2013-08-26 & 138.9--169.6              & 6h & 0.81\% & ---\\
GLEAM & 2013-11-05 & 72.3--133.8, 138.9--230.8 & 7h & 1.27\% & 0\% \\
GLEAM & 2013-11-25 & 72.3--133.8, 138.9--230.8 & 7h & 0.69\% & 0\% \\
EoR high& 2014-02-05 & 167.0--197.7              & 6h & 0.54\% & 0\%\\
GLEAM & 2014-03-16 & 72.3--133.8, 138.9--230.8 & 7h & 0.79\% & 0\% \\
GLEAM & 2014-03-17 & 72.3--133.8, 138.9--230.8 & 7h & 1.64\% & 3\% \\
EoR high& 2014-04-10 & 167.0--197.7              & 6h & 0.68\% & 0\%\\
GLEAM & 2014-06-18 & 72.3--133.8, 138.9--230.8 & 7h & 0.98\% & 0\% \\
Orbcomm RFI test&2014-08-27& 131.2--161.9     & 2$\times$8 m& 1.85\% & --- \\
\hline
\multicolumn{3}{|r|}{\textbf{Total}} & \textbf{63h} & \textbf{0.96\%} & \textbf{0.33\%}\\
\hline
\end{tabular}
\end{center}
\end{table*}

@@

\section{Detection results}
\noindent\begin{figure*}
\begin{center}\hspace*{-0.2cm}\includegraphics[width=18cm]{img/rfi-per-band}
\caption{RFI per subband}
\label{fig:rfi-per-band}
\end{center}
\end{figure*}

The fifth column of Table~\ref{tbl:obs-list} lists for each of the observation nights the ratio of samples that were classified as RFI by the \textsc{aoflagger}. This includes false detections, which are estimated in \citet{lofar-radio-environment} to account for approximately 0.4\% detections. Compared to the EoR observations, the GLEAM observations have a higher average of 1.05\% of RFI. This is mostly caused by the FM bands, which are only observed in GLEAM observations. The EoR low-band observation shows 0.81\% RFI, while the EoR high-band observations have an average of 0.58\%.

Fig.~\ref{fig:rfi-per-band} shows the overal detected RFI ratio per sub-band. The FM bands around 100~MHz and the Orbcomm bands around 138~MHz are clearly present in the data.

@@

\noindent\begin{figure*}
\subcaptionbox{RFI contamination found in the 2 m amateur band (146 MHz).}[\linewidth]{%
% obs data: 1061583576, 2013-08-26.
\includegraphics[width=9.1cm]{img/amateur-2m-band-example}\includegraphics[width=9.1cm]{img/amateur-2m-band-flagged}
\label{fig:amateur-2m}
}\\\vspace{3mm}%
\subcaptionbox{Short RFI burst of 2 s centred on 150.17 MHz. This shows that flagging at high time and frequency resolution is necessary to accurately detect RFI.}[\linewidth]{%
% obs data: 1061583576, 2013-08-26.
\includegraphics[width=9.1cm]{img/150_2_mhz_example}\includegraphics[width=9.1cm]{img/150_2_mhz_flagged}
\label{fig:150_2}
}\\\vspace{3mm}%
\subcaptionbox{A weakly-observed transmitter centred on 156.66 MHz. This same frequency is flagged in several correlations, and it can therefore be assumed this is real RFI. This is an example in which the \textsc{aoflagger} detects an event that would not be detected by regular thresholding. }[\linewidth]{%
% obs data: 1061583576, 2013-08-26.
\includegraphics[width=9.1cm]{img/156_7_mhz_example}\includegraphics[width=9.1cm]{img/156_7_mhz_flagged}
\label{fig:156_7}
}%
\caption{Examples of RFI events found in the EoR low band (138.9--197.7 MHz) in a 2-min snapshot with relatively high RFI contamination. These panels show the Stokes I amplitudes. In the right figures, the result of RFI detection is shown with purple. The horizontally flagged lines are flagged because they are 1.28-MHz subband edge or centre channels, which are unusable because of aliasing of the poly-phase filter bank and DC offsets, respectively.}
\end{figure*}

\noindent\begin{figure}
% obs data: 1061583576, 2013-08-26.
\begin{center}\hspace*{-0.2cm}\includegraphics[width=9.1cm]{img/dvb_example}
\caption{RFI from Digital Video Broadcasting (DVB). Because of its broadband nature, this kind of RFI is not well flagged in the initial \textsc{aoflagger} step, but is detectable in the global statistics. This type of RFI makes most of the 30.72~MHz bandwidth unusable. It is visible for about 2 h in 1 out of the 6 nights that observed this band. }
\label{fig:dvb_example}
\end{center}
\end{figure}

@@
